---
title : "Create Circos graph for CellPellet <> Supernatant Ligand-Receptor communication"
Author: Maria Kondili
Date: "03 May 2024"
header-includes:
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines=TRUE,commandchars=\\\{\}}
output: 
  html_document :
  toc: true
linestretch: 1.3
highlight: tango
df_print: paged
editor_options: 
  chunk_output_type: console
---


```{r, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,      # no code will be shown in html report
                      eval = TRUE,
                      fig.dim = c(10,8), # for larger space for a graph
                      pillar.sigfig = 5)  # number of decimals to show from a tibble

```


```{r libs}

suppressPackageStartupMessages(library(ggraph))
suppressPackageStartupMessages(library(igraph))
suppressPackageStartupMessages(library(tidyverse))
library(RColorBrewer)
suppressPackageStartupMessages(library(circlize))
suppressPackageStartupMessages(library(viridis))
library(gridBase)
suppressPackageStartupMessages(library(ComplexHeatmap))
#library(chorddiag)  # BiocManager::install("chorddiag")| devtools::install_github("mattflor/chorddiag")
```


```{r home_dir}

#setwd("/shared/projects/secretome/Project_OPMD_DMD/NicheNet")
work_dir <- "~/MYO_WORKSPACE/PROJECTS/Secretome/Secretome_DMD_OPMD/NicheNet/"

```

### Creating a table with strongest L-R interactions -> for Circos

Extract the pairs of ligands-receptors from nichenet results

```{r extract_strong_LR_function}

extract_LR_for_circos <- function(nichenet_res_path, celltype, cutoff) {
  
  if ( str_ends(string = nichenet_res_path,pattern = ".xlsx") )
    res_nn <- readxl::read_xlsx(nichenet_res_path, col_names = T)
  else if (str_ends(nichenet_res_path, pattern =".txt|.tsv|.csv") ) {
      res_nn <- read_delim(nichenet_res_path, delim="\t",col_names=TRUE,show_col_types = FALSE)
  }
  
  res_nn <- as.data.frame(res_nn)

  # rownames(res_nn) <- res_nn$ligands  #---> sometimes duplicate ligands,not possible rownames! 
  # res_nn <- res_nn %>% dplyr::select(-"ligands")
  # 
  LR_df <- data.frame(Ligand = character(), 
                      Receptor = character(), 
                      Interact_value = numeric(), 
                      stringsAsFactors = FALSE)  # Initialize an empty list to store results
  
  # Iterate through each row of the matrix
  for (i in 1:nrow(res_nn)) {
      # Find column indices where values exceed the cutoff in the current row
      # res_nn[,-1] : to ignore the "ligands" column 
      col_indices <- which(res_nn[i, -1 ] > cutoff)
    
      # If there are values above the cutoff in the current row
      if (length(col_indices) > 0) {
          for (j in col_indices) {
              LR_df <- rbind(LR_df, data.frame(
                             ligand =  res_nn$ligands[i],
                             receptor = colnames(res_nn)[j+1], 
                             interact_value  = res_nn[i, j+1],
                             cell_type=celltype )
                            )
          }
        # j+1 :to count for the ligands column that is not counted 
    }
  }
  
  return(LR_df)
}

```

> Call function :: All-Supernatant -vs- All-CellPellet

```{r all_vs_all_LRs_fun}

cat("\n\n~~~> This run will be based on OPMD Ligands-Receptors Results !!!\n\n ")

out_dir <- paste0(work_dir,"Outputs/OPMD/")

all_vs_all_file <- paste0(out_dir,"Ligands_Receptors_table_All_Supernatant_vs_All_CellPellet_proteins.xlsx")

cat("\nFile to be read :Ligands_Receptors_table_All_Supernatant_vs_All_CellPellet_proteins.xlsx ")

LR_all_vs_all_df <- extract_LR_for_circos(all_vs_all_file, celltype="CellPellet", cutoff = 0.4) 
  

```

> Call function :: All-Supernatant -vs- sc.Adipocytes 

```{r all-vs-adipo_LRs_fun}

all_vs_adipo_file <- paste0(out_dir,"Table_All_Supernatant=Ligands_vs_scAdipocytes=Receptors.txt")

cat("\nFile to be read :Table_All_Supernatant=Ligands_vs_scAdipocytes=Receptors.txt")

LR_all_vs_adipo_df <- extract_LR_for_circos( all_vs_adipo_file, celltype="Adipocytes",cutoff = 0.4 ) 


```

> Call function :: All Supernatant -vs- sc.Fibroblasts 

```{r all-vs-fibro_LRs_fun}

all_vs_fibro_file <- paste0(out_dir,"Table_All_Supernatant=Ligands_vs_scFibroblasts=Receptors.txt")

cat("\nFile to be read :Table_All_Supernatant=Ligands_vs_scFibroblasts=Receptors.txt")

LR_all_vs_fibro_df <- extract_LR_for_circos(all_vs_fibro_file, celltype="Fibroblasts", cutoff = 0.4 ) 

```


> Call function :: All Supernatant -vs- sc.Macrophages

```{r all-vs-macro_LRs_fun}

all_vs_macroph_file <- paste0(out_dir,"Table_All_Supernatant=Ligands_vs_scMacrophages=Receptors.txt")

cat("\nFile to be read :Table_All_Supernatant=Ligands_vs_scMacrophages=Receptors.txt")

LR_all_vs_macroph_df <- extract_LR_for_circos(all_vs_macroph_file, celltype="Macrophages", cutoff = 0.4 ) 

```

### Merge All L-R pairs in one table

Source: https://jokergoo.github.io/circlize_book/book/advanced-usage-of-chorddiagram.html 
<br>
(Figure : 15.12)

cat("\n~~~NOTE~~~")
cat("\nIn order to have a medium number of Ligands-Receptors to visualise,\n
    the LR pairs are sorted in decreasing order of interaction value,\n
    and the top 15 strongest will be taken from each NicheNet table.")


```{r merge_LRs_tables}

nichenet_LR <- dplyr::bind_rows(LR_all_vs_all_df ,
                                LR_all_vs_adipo_df, 
                                LR_all_vs_fibro_df,
                                LR_all_vs_macroph_df  )

## Choose the 4 strongest LR-interactions per pair/cell-type:

keep_top10_highInteractions_LR <- function(LR_df) {
  
    maxI_LR <- LR_df[order(LR_df$interact_value, decreasing = T),][1:15,] 
    ## remove one line of the ligand being also a receptor
    if ( any(maxI_LR$ligand %in% maxI_LR$receptor)) { 
        to_rmv <- which(maxI_LR$ligand %in% maxI_LR$receptor)
        # All pairs removed ! BUT: have to keep the pair once ...! 
        maxI_LR <- maxI_LR[-to_rmv,]
      }
    return(maxI_LR)
}

##> apply function for each data.frame of LR
maxI_LR_list <- map(list("cp"=LR_all_vs_all_df ,
                         "adipo"=LR_all_vs_adipo_df,
                         "fibro"=LR_all_vs_fibro_df,
                         "macroph"=LR_all_vs_macroph_df),
                  ~keep_top10_highInteractions_LR(.x))


mini_nn_LR <- dplyr::bind_rows(maxI_LR_list)

print(mini_nn_LR)
```

Remove again Receptors that are also Ligands, from different cell-types
(not seen in the function)

```{r filter_Ligands_ovl_Receptors }

to_rmv <- which( mini_nn_LR$receptor %in% mini_nn_LR$ligand )
mini_nn_LR <- mini_nn_LR[ -to_rmv, ]
```

Verify that Receptors are unique :

```{r unique_receptors }

mini_nn_LR <- mini_nn_LR[-which(duplicated(mini_nn_LR$receptor)),]

mini_nn_LR
```

Create table of Adjacency Ligands-Receptors 

```{r  define_cell_types_and_LR_mat }

celltypes <- mini_nn_LR$cell_type
names(celltypes) <- mini_nn_LR$receptor


## Create table of frequency of an interaction of Ligand with a Receptor 
LR_mat <- with(mini_nn_LR , table(ligand,receptor))


#!ATTENTION!
# LR_mat -> colnames ="Receptors" are in Alphab.order, not as in data.frame !
# This is why we 'll Match Receptors names --> reorder according to data.frame
matched_ids <- match( mini_nn_LR$receptor, colnames(LR_mat) )
# 
# ##> change order of columns :
LR_mat <- LR_mat[,matched_ids]

#cat("\nNow LR_mat has columns with same order as mini_nn_LR.\n")
#colnames(LR_mat)
#mini_nn_LR$receptor
```

Assign Genes to Cell-types Groups

```{r prep_circos_info }

groups <- structure( c( rep("Ligands", nrow(LR_mat)), 
                        celltypes ), 
                    names = unlist(dimnames(LR_mat)) )

groups_df <- data.frame( "groups" = groups )
print(groups_df)

glue::glue("\nThe data contain {length(celltypes)}  receptors from different sources," )
glue::glue("\n...and {nrow(LR_mat)} ligands" )
glue::glue("..giving a total of {length(celltypes) + nrow(LR_mat)} features to participate in the Circos Plot." )


cells_groups <- levels(as.factor(celltypes))
#bind_cols(groups, names(groups) ) %>% print(n= ncol(LR_mat)+nrow(LR_mat) )
cells_groups

```

Define the colours that the label-bars will get

```{r prep_colour_labels}


#(ATTENTION : Add as many colours as cells_groups )

groups_df <- groups_df %>% mutate("grid_colour"= c(rep("darkorchid3", length(which(groups =="Ligands"))), 
                                                  rep("chocolate4",length(which(groups == "CellPellet" ))),    
                                                  rep("orange2",length(which(groups == "Adipocytes" ))),          
                                                  rep("aquamarine3",length(which(groups == "Macrophages" ))))
)

print(groups_df)

grid_col <- groups_df$grid_colour
names(grid_col) <- rownames(groups_df)

```

### Circos 1 :: Classic Chords with random colours (based on frequency of interaction ={0,1})

```{r create_circos }

cat("\nNumber of ligands & Receptors is equal the number of colours given to each: \n")
length(unlist(dimnames(LR_mat))) == length(groups_df$grid_colour)

circos.clear()
circos.par(canvas.ylim=c(-1, 1), # edit  canvas size 
           track.margin = c(0.01, 0.01)) # adjust bottom and top margin


##> Chords with Random Colours
chordDiagram(LR_mat,
             transparency=0.5,
             big.gap = 10,
             group = groups,
             grid.col = grid_col, #only for small-genes-boxes colour
             annotationTrack = "grid",
             annotationTrackHeight = c(0.08,0.08), ##--> gives thickness on bars
             link.target.prop = TRUE,
             preAllocateTracks = list(
                track.height = mm_h(3),
                track.margin = c(mm_h(3),0)),
                col=rand_color(length(LR_mat)) ) #-> colors of chords=random,
                                                 #  length(LR_mat) = number of all L,R.

# + Arrows:  directional=1,  direction.type = "arrows",

##> Add Receptors_Names in "Track"=2 ( internal circle)
circos.track(track.index=2, panel.fun = function(x,y) { 
               sector.index = get.cell.meta.data("sector.index")
               xlim = get.cell.meta.data("xlim")
               ylim = get.cell.meta.data("ylim")
               circos.text(mean(xlim), mean(ylim), 
                           sector.index, cex = 0.4,
                           niceFacing = TRUE,col = "white")
            }, bg.border = NA )



### 2nd Layer of Labels :: Cell-types (Source of genes )


# Ligands layer
highlight.sector(rownames(LR_mat), track.index = 1, col = unique(groups_df$grid_colour)[1], 
    text = unique(groups_df$groups)[1], cex = 0.8, lwd = 1.5, text.col = "white", niceFacing = TRUE)


# CellPellet
highlight.sector(colnames(LR_mat)[which(mini_nn_LR$cell_type == unique(groups_df$groups)[2] )], #cell_group[1]
                 track.index = 1, col = unique(groups_df$grid_colour)[2], 
                 text =unique(groups_df$groups)[2] , cex = 0.8, lwd = 1.5,
                 text.col = "white", niceFacing = TRUE )

# Adipocytes
highlight.sector(colnames(LR_mat)[which(mini_nn_LR$cell_type == unique(groups_df$groups)[3])], #cell_group[2]
                 track.index = 1, col = unique(groups_df$grid_colour)[3] , 
                 text = unique(groups_df$groups)[3] , cex = 0.8, lwd = 1.5, 
                 text.col = "white", niceFacing = TRUE)


# Macrophages
highlight.sector(colnames(LR_mat)[which(mini_nn_LR$cell_type == unique(groups_df$groups)[4])],   #cell_group[3]
                 track.index = 1, col = unique(groups_df$grid_colour)[4], 
                 text = unique(groups_df$groups)[4], cex = 0.8, lwd = 1.5, 
                 text.col = "white",  text.vjust = 0.5,niceFacing = TRUE )


```

### Circos 2 :: Chords colored according to Interaction Value Intensity (from nichenet)

NOTE: LR_mat must contain the interaction value in each cell, instead of the {0,1}

#### Create table of Interactions of Ligands with Receptors 

```{r prepare_LR_mat_with_Interaction }

#instead of : LR_mat <- with(mini_nn_LR , table(ligand,receptor))
#! Attention: Must remove "cell_type" column ,so that ligands with multiple receptors are merged in 1 line

LR_mat.i <- mini_nn_LR %>% dplyr::select(-c("cell_type")) %>% 
                            tidyr::pivot_wider(names_from = receptor, 
                                               values_from="interact_value",
                                               values_fill = 0)
## Remove "ligand" column -> keep as rownames
LR_mat.i <- as.data.frame(LR_mat.i)
rownames(LR_mat.i) <- LR_mat.i$ligand
LR_mat.i <- LR_mat.i %>% dplyr::select(-"ligand")

print(LR_mat.i)

```

#### Create ChordDiagram

```{r create_circos_2 }


circos.par(canvas.ylim=c(-1, 1), # edit  canvas size 
           track.margin = c(0.01, 0.01)) # adjust bottom and top margin


col_values <- c(round(min(mini_nn_LR$interact_value),1), 
                round(max(mini_nn_LR$interact_value),1) )

col_fun <- circlize::colorRamp2(col_values, c("blue", "red"))

LR_mat.i <- as.matrix(LR_mat.i) 

chordDiagram(LR_mat.i,
             transparency=0.5,
             big.gap = 10,
             group = groups,
             grid.col = grid_col,        #only for small-genes-boxes colour
             annotationTrack = "grid",
             annotationTrackHeight = c(0.08,0.08), ##--> gives thickness on bars
             link.target.prop = TRUE,
             directional=1,         #shows little bars on Ligands showing to which R is connected
             preAllocateTracks = list(
                                track.height = mm_h(3),
                                track.margin = c(mm_h(3),0)),
             col=function(x) col_fun(x)) 
            


```

#### Add Legend and Colour intensity according to Interaction Value of L-Rs

#> Doc : https://jokergoo.github.io/ComplexHeatmap-reference/book/legends.html


```{r add_legend }

## Add a Legend for the color-variability of chords 
lgd <- ComplexHeatmap::Legend(at=col_values,
                              col_fun = col_fun, 
                              title = "Interaction")

ComplexHeatmap::draw(lgd, x = unit(1, "cm"), y = unit(0, "cm"),just=c("left","bottom") )

cat("\n\n\n~~>Legend shows colour-code of chords according to Interaction Value: blue = min, red = max ")

##> Add Receptors_Names in "Track"=2 ( internal circle)
circos.track(track.index=2, panel.fun = function(x,y) { 
               sector.index = get.cell.meta.data("sector.index")
               xlim = get.cell.meta.data("xlim")
               ylim = get.cell.meta.data("ylim")
               circos.text(mean(xlim), mean(ylim), 
                           sector.index, cex = 0.6,
                           niceFacing = TRUE,col = "white")
            }, bg.border = NA )


### 2nd Layer of Labels :: Cell-types (Source of genes )

## Parameters to change : 
## 1. col = unique(groups_df$grid_colour)[1] -> [2], [3],[4] according to nb.of cell-types
## 2. text = unique(groups_df$groups)[1] & sector.index = [which(mini_nn_LR$cell_type ==...]


# Ligands layer
highlight.sector(rownames(LR_mat.i),
                 track.index = 1, 
                 col = unique(groups_df$grid_colour)[1], 
                 text = unique(groups_df$groups)[1], 
                 cex = 0.8, lwd = 1.5, text.col = "white", 
                 niceFacing = TRUE )



# CellPellet
highlight.sector(colnames(LR_mat.i)[which(mini_nn_LR$cell_type == unique(groups_df$groups)[2] )], #cell_group[1]
                 track.index = 1, 
                 col = unique(groups_df$grid_colour)[2], 
                 text = unique(groups_df$groups)[2] , 
                 cex = 0.8, lwd = 1.5, text.col = "white", 
                 niceFacing = TRUE )


# Adipocytes
highlight.sector(colnames(LR_mat.i)[which(mini_nn_LR$cell_type == unique(groups_df$groups)[3])], #cell_group[2]
                 track.index = 1, 
                 col = unique(groups_df$grid_colour)[3] , 
                 text = unique(groups_df$groups)[3], 
                 cex = 0.8, lwd = 1.5, 
                 text.col = "white", 
                 niceFacing = TRUE)


# Macrophages
highlight.sector(colnames(LR_mat.i)[which(mini_nn_LR$cell_type == unique(groups_df$groups)[4])],   #cell_group[3]
                 track.index = 1, 
                 col = unique(groups_df$grid_colour)[4], 
                 text = unique(groups_df$groups)[4], 
                 cex = 0.8, lwd = 1.5, 
                 text.col = "white",  
                 text.vjust = 0.5,
                 niceFacing = TRUE )


circos.clear()

```

### Create ChordDiagram B' : only with Ligands and Cell-Pellet (10/10/24)

```{r nichenet_LR_for_cellpellet, eval=F}

#nichenet.cp <- nichenet_LR %>% subset(cell_type=="CellPellet")

mini_nn_LR.cp <- mini_nn_LR %>% subset(cell_type=="CellPellet")

cp_celltype <- mini_nn_LR.cp$cell_type
names(cp_celltype) <- mini_nn_LR.cp$receptor


## Create table of frequency of an interaction of Ligand with a Receptor 
LR_mat.cp <- with(mini_nn_LR.cp, table(ligand,receptor))

print(LR_mat.cp)

#!ATTENTION!
# LR_mat -> colnames ="Receptors" are in Alphab.order, not as in data.frame !
# This is why we 'll Match Receptors names --> reorder according to data.frame
matched_ids <- match( mini_nn_LR.cp$receptor, colnames(LR_mat.cp) )

# ##> change order of columns :
LR_mat.cp <- LR_mat.cp[,matched_ids]
groups_df.cp <- subset(groups_df, groups %in% c("Ligands","CellPellet"))


groups_df.cp <- groups_df.cp %>% 
                mutate("grid_colour"= c(rep("darkorchid3", length(which(groups=="Ligands"))), 
                                        rep("darkslategray",length(which(groups == "CellPellet")))   
                                      ))



grid_col.cp <- groups_df.cp$grid_colour
names(grid_col.cp) <- rownames(groups_df.cp)

```


```{r circos_only_with_cellpellet_graph, eval=F}

circos.par(canvas.ylim=c(-1, 1), # edit  canvas size 
           track.margin = c(0.01, 0.01)) # adjust bottom and top margin


col_values.cp <- c(round(min(mini_nn_LR.cp$interact_value),1), 
                round(max(mini_nn_LR.cp$interact_value),1) )

col_fun <- circlize::colorRamp2(col_values.cp, c("blue", "red"))

glue::glue("\n\n~~> Colours will be set on following values of Interaction :min={col_values[1]}, max={col_values[2]}")


chordDiagram(LR_mat.cp,
             transparency=0.5,
             big.gap = 10,
             group = groups,
             grid.col = grid_col.cp,        #only for small-genes-boxes colour
             annotationTrack = "grid",
             annotationTrackHeight = c(0.08,0.08), ##--> gives thickness on bars
             link.target.prop = TRUE,
             directional=1,         #shows little bars on Ligands showing to which R is connected
             preAllocateTracks = list(
                                track.height = mm_h(3),
                                track.margin = c(mm_h(3),0)),
             col=function(x) col_fun(x)) 
            

#### Add Legend and Colour intensity according to Interaction Value of L-Rs

#> Doc : https://jokergoo.github.io/ComplexHeatmap-reference/book/legends.html


## Add a Legend for the color-variability of chords 
lgd <- ComplexHeatmap::Legend(at=col_values.cp,
                              col_fun = col_fun, 
                              title = "Interaction")

ComplexHeatmap::draw(lgd, x = unit(1, "cm"), y = unit(0, "cm"),just=c("left","bottom") )

cat("\n\n\n~~>Legend shows colour-code of chords according to Interaction Value: blue = min, red = max ")


##> Add Receptors_Names in "Track"=2 ( internal circle)
circos.track(track.index=2, panel.fun = function(x,y) { 
               sector.index = get.cell.meta.data("sector.index")
               xlim = get.cell.meta.data("xlim")
               ylim = get.cell.meta.data("ylim")
               circos.text(mean(xlim), mean(ylim), 
                           sector.index, cex = 0.6,
                           niceFacing = TRUE,col = "white")
            }, bg.border = NA )


#### 2nd Layer of Labels :: Cell-types (Source of genes )


# Ligands layer
highlight.sector(rownames(LR_mat.cp),
                 track.index = 1, 
                 col = unique(groups_df.cp$grid_colour)[1], 
                 text = unique(groups_df.cp$groups)[1], 
                 cex = 0.8, lwd = 1.5, text.col = "white", 
                 niceFacing = TRUE )



# CellPellet
highlight.sector(colnames(LR_mat.cp)[which(mini_nn_LR.cp$cell_type == unique(groups_df.cp$groups)[2])], #cell_group[1]
                 track.index = 1, 
                 col = unique(groups_df.cp$grid_colour)[2], 
                 text = unique(groups_df.cp$groups)[2] , 
                 cex = 0.8, lwd = 1.5, text.col = "white", 
                 niceFacing = TRUE )


```

### Create ChordDiagram C' : only with POI-Ligands-OPMD and All Cell-Pellet (updated October 2024)

```{r nichenet_LR_poi_Ligands}

poi_sn_vs_all_cp_file <- paste0(out_dir,"Table_POI_Supernatant=Ligands_vs_All_CellPellet=Receptors.xlsx")
glue::glue("\nFile to be read : {basename(poi_sn_vs_all_cp_file)}")

LR_poiSN_vs_CP_df <- extract_LR_for_circos(poi_sn_vs_all_cp_file, celltype="CellPellet", cutoff=0.3 ) 
LR_poiSN_vs_CP_df %>% head

```


```{r prepare_LR_mat_poi_Supernat_ligands}

to_rmv <- which( LR_poiSN_vs_CP_df$receptor %in% LR_poiSN_vs_CP_df$ligand )
to_rmv

LR_poiSN_vs_CP_df  <- LR_poiSN_vs_CP_df[-which(duplicated(LR_poiSN_vs_CP_df$receptor)),]

###
### adjacency table 
###

celltypes <- LR_poiSN_vs_CP_df$cell_type
names(celltypes) <- LR_poiSN_vs_CP_df$receptor

LR_mat.poi.sn <- with(LR_poiSN_vs_CP_df , table(ligand,receptor))

print(LR_mat.poi.sn)

matched_ids <- match( LR_poiSN_vs_CP_df$receptor, colnames(LR_mat.poi.sn) )
#> change order of columns :
LR_mat.poi.sn <- LR_mat.poi.sn[,matched_ids]


### groups 

groups.poi.sn <- structure( c( rep("Ligands", nrow(LR_mat.poi.sn)), celltypes), 
                              names = unlist(dimnames(LR_mat.poi.sn)) )

groups_df.poi.sn <- data.frame( "groups" = groups.poi.sn )
print(groups_df.poi.sn)

groups_df.poi.sn <- groups_df.poi.sn %>% 
                    mutate("grid_colour"= c(rep("firebrick3", length(which(groups == "CellPellet" ))), 
                                            rep("mediumpurple3",length(which(groups == "Ligands" )))))   
                                                 
print(groups_df.poi.sn)

### grid color 

grid_col.poi.sn <- groups_df.poi.sn$grid_colour
names(grid_col.poi.sn) <- rownames(groups_df.poi.sn)


```

Circos Plot 

(LR_poiSN_vs_CP_df <- mini_nn_LR)

```{r circos_with_poi_Ligands_vs_All_CP}

circos.par(canvas.ylim=c(-1, 1), # edit  canvas size 
           track.margin = c(0.01, 0.01)) # adjust bottom and top margin


col_values.poi.sn <- c(round(min(LR_poiSN_vs_CP_df$interact_value),1), 
                round(max(LR_poiSN_vs_CP_df$interact_value),1) )

col_fun.poi.sn <- circlize::colorRamp2(col_values.poi.sn, c("blue", "red"))
col_fun.poi.sn

chordDiagram(LR_poiSN_vs_CP_df,
             transparency=0.5,
             big.gap = 10,
             group = groups.poi.sn,
             grid.col = grid_col.poi.sn,        #only for small-genes-boxes colour
             annotationTrack = "grid",
             annotationTrackHeight = c(0.08,0.08), ##--> gives thickness on bars
             link.target.prop = TRUE,
             directional=1,         #shows little bars on Ligands showing to which R is connected
             preAllocateTracks = list(
                                track.height = mm_h(3),
                                track.margin = c(mm_h(3),0)),
             col=function(x) col_fun(x)) 
            

#### Add Legend and Colour intensity according to Interaction Value of L-Rs

#> Doc : https://jokergoo.github.io/ComplexHeatmap-reference/book/legends.html


## Add a Legend for the color-variability of chords 
lgd <- ComplexHeatmap::Legend(at=col_values.poi.sn,
                              col_fun = col_fun, 
                              title = "Interaction")

ComplexHeatmap::draw(lgd, x = unit(1, "cm"), y = unit(0, "cm"),just=c("left","bottom") )

cat("\n\n\n~~>Legend shows colour-code of chords according to Interaction Value: blue = min, red = max ")


##> Add Receptors_Names in "Track"=2 ( internal circle)
circos.track(track.index=2, panel.fun = function(x,y) { 
               sector.index = get.cell.meta.data("sector.index")
               xlim = get.cell.meta.data("xlim")
               ylim = get.cell.meta.data("ylim")
               circos.text(mean(xlim), mean(ylim), 
                           sector.index, cex = 0.6,
                           niceFacing = TRUE,col = "white")
            }, bg.border = NA )


#### 2nd Layer of Labels :: Cell-types => Cell Pellet (Source of genes )


# Ligands layer
highlight.sector(LR_poiSN_vs_CP_df$ligand,
                 track.index = 1, 
                 col = unique(groups_df.poi.sn$grid_colour)[1], 
                 text = unique(groups_df.poi.sn$groups)[1], 
                 cex = 0.8, lwd = 1.5, text.col = "white", 
                 niceFacing = TRUE )



# CellPellet
highlight.sector(LR_poiSN_vs_CP_df$receptor[which(LR_poiSN_vs_CP_df$cell_type == unique(groups_df.poi.sn$groups)[2])], #cell_group[1]
                 track.index = 1, 
                 col = unique(groups_df.poi.sn$grid_colour)[2], 
                 text = unique(groups_df.poi.sn$groups)[2] , 
                 cex = 0.8, lwd = 1.5, text.col = "white", 
                 niceFacing = TRUE )


```

### Create ChordDiagram D' : only with POI-FAPs(in_vitro) vs sc.MuSC (DeMicheli) [ 15/10/24 ]

```{r nichenet_LR_poi_FAPs}

# poi_faps_vs_musc <- paste0(out_dir,"../POI_FAPs_invitro/NicheNet_POI-FAPs_inVitro=Ligands_vs_scDeMicheli_MuSC=Receptors.xlsx")

## or, for Upregulated POI-FAPs:
poi_faps_vs_musc <- paste0(out_dir,"../POI_FAPs_invitro/",
                           "NicheNet_Upregul_POI-FAPs_inVitro=Ligands_vs_scDeMicheli_MuSC=Receptors.xlsx")

glue::glue("\nFile to be read : {basename(poi_faps_vs_musc)}")

LR_poi_faps_df <- extract_LR_for_circos(poi_faps_vs_musc, celltype="MuSC", cutoff = 0.2) 
LR_poi_faps_df %>% head
LR_poi_faps_df %>% dim

```


```{r prepare_LR_mat_FAPs-MuSC}

## remove Receptors that are also Ligands. Interferes with colouring to a protein in Circos.

to_rmv <- which( LR_poi_faps_df$receptor %in% LR_poi_faps_df$ligand )
#show line with receptor to rmv:
print(LR_poi_faps_df[to_rmv,])
# show other lines with ligand same as the receptor-2-remove:
LR_poi_faps_df[ which(LR_poi_faps_df$ligand == LR_poi_faps_df$receptor[to_rmv]),]


if (length(to_rmv)>0) {
  LR_poi_faps_df <- LR_poi_faps_df[-to_rmv,]
}

## remove Duplicate Receptors, as it is confounding when making Circos, to have them in different positions! 
## One Receptor from 2 Ligands is not possible,due to duplicate col-name of the table !
LR_poi_faps_df  <- LR_poi_faps_df[-which(duplicated(LR_poi_faps_df$receptor)),]


###
### adjacency table 
###

celltypes <- LR_poi_faps_df$cell_type
names(celltypes) <- LR_poi_faps_df$receptor

LR_mat.poi.faps <- with(LR_poi_faps_df , table(ligand,receptor))

print(LR_mat.poi.faps)

matched_ids <- match( LR_poi_faps_df$receptor, colnames(LR_mat.poi.faps))
#> change order of columns :
LR_mat.poi.faps <- LR_mat.poi.faps[,matched_ids]


###
### Groups 
###

##> 'groups' is a structure (vector) with protein-names and their group(ligand,receptor,celltype)
##> based on this we ll assign the colours 
##> rep("Upregul_FAPs" or "Ligands_FAPs",nrow(LR_mat.poi.faps) )

groups.poi.faps <- structure(c(rep("Upregul_FAPs", nrow(LR_mat.poi.faps)), celltypes), 
                              names = unlist(dimnames(LR_mat.poi.faps)))


groups_df.poi.faps <- data.frame( "groups" = groups.poi.faps )
print(groups_df.poi.faps)

groups_df.poi.faps <- groups_df.poi.faps %>% 
                      mutate("grid_colour"= c( rep("mediumpurple3",length(which(groups =="Upregul_FAPs"))),
                                               rep("firebrick3", length(which(groups == "MuSC")))))   
                                                 
print(groups_df.poi.faps)

length(unlist(dimnames(LR_mat.poi.faps))) == length(groups_df.poi.faps$grid_colour)


### grid color 

grid_col.poi.faps <- groups_df.poi.faps$grid_colour
names(grid_col.poi.faps) <- rownames(groups_df.poi.faps)

grid_col.poi.faps

```

Circos Plot 

(LR_poi_faps_df <- mini_nn_LR)

```{r circos_with_poi_FAPs_vs_MuSC }

#circos.clear()
circos.par(canvas.ylim=c(-1, 1), # edit  canvas size 
           track.margin = c(0.01, 0.01)) # adjust bottom and top margin


col_values.poi.faps <- c(round(min(LR_poi_faps_df$interact_value),1), 
                         round(max(LR_poi_faps_df$interact_value),1) )

col_fun.poi.faps <- circlize::colorRamp2(col_values.poi.faps, c("blue", "red"))

chordDiagram(LR_poi_faps_df,
             transparency=0.5,
             big.gap = 10,
             group = groups.poi.faps,
             grid.col = grid_col.poi.faps,         # only for small-genes-boxes colour
             annotationTrack = "grid",
             annotationTrackHeight = c(0.08,0.08), ##--> gives thickness on bars
             link.target.prop = TRUE,
             directional=1,         #shows little bars on Ligands showing to which R is connected
             preAllocateTracks = list(
                                track.height = mm_h(3),
                                track.margin = c(mm_h(3),0)),
             col=function(x) col_fun(x)) 
            

#### Add Legend and Colour intensity according to Interaction Value of L-Rs

#> Doc : https://jokergoo.github.io/ComplexHeatmap-reference/book/legends.html


## Add a Legend for the color-variability of chords 
lgd <- ComplexHeatmap::Legend(at=col_values.poi.faps,
                              col_fun = col_fun, 
                              title = "Interaction")

ComplexHeatmap::draw(lgd, x = unit(1, "cm"), y = unit(0, "cm"), just=c("left","bottom"))

cat("\n\n\n~~>Legend shows colour-code of chords according to Interaction Value: blue = min, red = max ")


##> Add Receptors_Names in "Track"=2 ( internal circle)
circos.track(track.index=2, panel.fun = function(x,y) { 
               sector.index = get.cell.meta.data("sector.index")
               xlim = get.cell.meta.data("xlim")
               ylim = get.cell.meta.data("ylim")
               circos.text(mean(xlim), mean(ylim), 
                           sector.index, cex = 0.5,
                           niceFacing = TRUE,col = "white")
            }, bg.border = NA )


#### 2nd Layer of Labels :: Cell-types => Cell Pellet (Source of genes )


# Ligands layer
highlight.sector(LR_poi_faps_df$ligand,
                 track.index = 1, 
                 col = unique(groups_df.poi.faps$grid_colour)[1], 
                 text = unique(groups_df.poi.faps$groups)[1], 
                 cex = 0.8, lwd = 1.5, text.col = "white", 
                 niceFacing = TRUE )



# MuSC layer
highlight.sector(LR_poi_faps_df$receptor[which(LR_poi_faps_df$cell_type == unique(groups_df.poi.faps$groups)[2])], #cell_group[1]
                 track.index = 1, 
                 col = unique(groups_df.poi.faps$grid_colour)[2], 
                 text = unique(groups_df.poi.faps$groups)[2] , 
                 cex = 0.8, lwd = 1.5, text.col = "white", 
                 niceFacing = TRUE )


#circos.clear()

```

```{r session_info}

sessionInfo()

```
